The Evolution of Phenotypically Invariant Genetic Networks

### 2017-01-31 kalman_walk.tex
assigned tasks

\documentclass[11 pt]{article}
\usepackage{amsmath, amssymb, color, xcolor}
\usepackage{graphicx}
\usepackage{fullpage}
\usepackage[hidelinks]{hyperref}
\usepackage{tikz}
\usetikzlibrary{arrows.meta, shapes}
\usepackage{caption, subcaption}
\usepackage{natbib} % gives us \citet: Author (year) and \citep: (Author; year)

\newcommand{\plr}[1]{{\color{blue}\it #1}}
\newcommand{\jss}[1]{{\color{olive}\it #1}}
% \newcommand{\ddt}{\frac{d}{dt}}
\newcommand{\ddt}{\dot}
\newcommand{\ro}{{ro}}
\newcommand{\nro}{{\bar{r}o}}
\newcommand{\rno}{{r\bar{o}}}
\newcommand{\nrno}{{\bar{r}\bar{o}}}
\newcommand{\reachable}{\mathcal{R}}
\newcommand{\unobservable}{\bar{\mathcal{O}}}
\newcommand{\R}{\mathbb{R}}
\newcommand{\E}{\mathbb{E}}
\renewcommand{\P}{\mathbb{P}}

@@

\begin{enumerate}
    \item Add biology to introduction. (keep an eye out for concrete numbers we can use!)

    \item Do some order-of-magnitude calculation to figure out how fast systems should drift.

    \item Write out simple examples in biological words.

    \item Rephrase hybrid inviability example in biological terms.  Discuss F1's (or heterozygotes) versus F2's.

    \item Clarify our assumptions around $x(0)$.

    \item Note we don't include $D$ for simplicity but don't.

    \item Figure out how to specify $P$ for Kalman decomposition.

    \item Prove or find citation for uniqueness of minimal system up to basis change.

    \item Work examples (minimal; not minimal; with/without B,C fixed) of Kalman decomposition and description of moving along the walk.

    \item Write down and decide between options for parameterizing the random walk on the neutral manifold:
        use Kalman decomposition, and parameterize Brownian motion so that it has constant speed;
        OR consider Josh's parameteriziation like this;
        OR just say we do Brownian motion on the manifold (more or less implicit characterization).

@@

    \item Describe model leading to constraints on regulation so we have a stationary distribution.
        Ask questions of stationary distribution: robustness, connectivity, xxx?

@@

@@

\title{Mathematical Gene Network Evolution}
\maketitle

@@

    \begin{figure}
        \begin{center}
            \begin{tikzpicture}
            \begin{scope}[every node/.style={circle,thick,draw}]
                \node (A) at (0,0) {$x_{1}$};
                \node[dashed] (B) at (3,0) {$x_{2}$};
                \node[shape=rectangle] (U) at (1.5,2) {$u$};
                \node[shape=rectangle] (y) at (1.5,-2) {$y$};
            \end{scope}

@@

\begin{figure}
  \centering
  \begin{subfigure}{0.5\textwidth}
    \centering
\begin{tikzpicture}
\begin{scope}[every node/.style={circle,thick,draw}]
  \node (A) at (0,0) {$x_{1}$};
    \node[dashed] (B) at (0,-4) {$x_{2}$};
    \node[shape=rectangle, dashed] (U) at (3,-2) {$u(t)$};
\end{scope}

\begin{scope}[>={Stealth[black]},
              every node/.style={fill=white,circle},
              every edge/.style={draw=black, thick}]
    %\path [->] (A) edge[bend left] node {$c$} (B);
    \path [->] (B) edge node {\tiny $\gamma$} (A); 
    \path[->] (U) edge[dashed] node {\tiny $1$} (A);
    \path[->] (U) edge[dashed] node {\tiny $1$} (B);
\end{scope}
\begin{scope}[>={Stealth[black]},
              every edge/.style={draw=black, thick}]
    \path [->] (A) edge[loop left] node {\tiny $\lambda_{1}$} (A);
    \path [->] (B) edge[loop left] node {\tiny $\lambda_{2}$} (B);
\end{scope}

\end{tikzpicture}
    \caption{Gene Regulatory Network $A(0)$.}
  \end{subfigure}%
  \begin{subfigure}{0.5\textwidth}
    \centering
\begin{tikzpicture}
\begin{scope}[every node/.style={circle,thick,draw}]
  \node (A) at (0,0) {$x_{1}$};
    \node[dashed] (B) at (0,-4) {$x_{2}$};
    \node[shape=rectangle, dashed] (U) at (3,-2) {$u(t)$};
\end{scope}

\begin{scope}[>={Stealth[black]},
              every node/.style={fill=white,circle},
              every edge/.style={draw=black, thick}]
    \path [->, sloped] (A) edge[bend left] node {\tiny $(\lambda_{2} - \lambda_{1} - \frac{\gamma}{2})$} (B);
    \path [->, sloped] (B) edge[bend left] node {\tiny $\frac{\gamma}{2}$} (A); 
    \path[->] (U) edge[dashed] node {\tiny $1$} (A);
    \path[->] (U) edge[dashed] node {\tiny $1$} (B);
\end{scope}
\begin{scope}[>={Stealth[black]},
              every edge/.style={draw=black, thick}]
    \path [->] (A) edge[loop left] node {\tiny $(\lambda_{1} + \frac{\gamma}{2})$} (A);
    \path [->] (B) edge[loop left] node {\tiny $(\lambda_{2}-\frac{\gamma}{2})$} (B);
\end{scope}

>>>>

\documentclass[11 pt]{article}
\usepackage{amsmath, amssymb, color, xcolor}
\usepackage{graphicx}
\usepackage{fullpage}
\usepackage[hidelinks]{hyperref}
\usepackage{tikz}
\usetikzlibrary{arrows.meta, shapes}
\usepackage{caption, subcaption}
\usepackage{natbib} % gives us \citet: Author (year) and \citep: (Author; year)
\usepackage{authblk}
\newcommand{\plr}[1]{{\color{blue}\it #1}}
\newcommand{\jss}[1]{{\color{olive}\it #1}}
% \newcommand{\ddt}{\frac{d}{dt}}
\newcommand{\ddt}{\dot}
\newcommand{\ro}{{ro}}
\newcommand{\nro}{{\bar{r}o}}
\newcommand{\rno}{{r\bar{o}}}
\newcommand{\nrno}{{\bar{r}\bar{o}}}
\newcommand{\reachable}{\mathcal{R}}
\newcommand{\unobservable}{\bar{\mathcal{O}}}
\newcommand{\R}{\mathbb{R}}
\newcommand{\E}{\mathbb{E}}
\renewcommand{\P}{\mathbb{P}}

@@

    \item \emph{(josh)} Add biology to introduction. (keep an eye out for concrete numbers we can use!)

    \item Do some order-of-magnitude calculation to figure out how fast systems should drift.
        \begin{itemize}
            \item \emph{(peter)} Work out the math for how to do this.
            \item \emph{(josh)} Find real parameters to put into this.
        \end{itemize}

    \item \emph{(josh)} Write out simple examples in biological words.

    \item \emph{(josh)} Rephrase hybrid inviability example in biological terms.  Discuss F1's (or heterozygotes) versus F2's.

    \item \emph{(josh)} Clarify our assumptions around $x(0)$.

    \item \emph{(josh)} Note we don't include $D$ for simplicity but don't.

    \item \emph{(peter)} Figure out how to specify $P$ for Kalman decomposition.

    \item \emph{(josh,peter)} Prove or find citation for uniqueness of minimal system up to basis change.

    \item Work examples (minimal; not minimal; with/without B,C fixed) of Kalman decomposition and description of moving along the walk.
        \begin{itemize}
            \item \emph{(josh)} Work out parameterization.
            \item \emph{(peter)} Describe Brownian motion.
        \end{itemize}

    \item \emph{(peter)} Write down and decide between options for parameterizing the random walk on the neutral manifold:
        use Kalman decomposition, and parameterize Brownian motion so that it has constant speed;
        OR consider Josh's parameteriziation like this;
        OR just say we do Brownian motion on the manifold (more or less implicit characterization).

    \item \emph{(peter)} Describe model leading to constraints on regulation so we have a stationary distribution.
        Ask questions of stationary distribution: robustness, connectivity, xxx?

@@

@@

    \item Simulate.

@@

\title{The Evolution of Phenotypically Invariant Genetic Networks}
\author[1]{Joshua S. Schiffman}
\author[1,2]{Peter L. Ralph}
\affil[1]{University of Southern California, Los Angeles, California}
\affil[2]{University of Oregon, Eugene, Oregon}
\affil[ ]{\texttt{jsschiff@usc.edu} \qquad \texttt{plr@uoregon.edu}}
\maketitle

@@

    \begin{figure}
        \begin{center}
            \begin{tikzpicture}
            \begin{scope}[every node/.style={circle,thick,draw}]
                \node (A) at (0,0) {$x_{1}$};
                \node[dashed] (B) at (3,0) {$x_{2}$};
                \node[shape=rectangle] (U) at (1.5,2) {input};
                \node[shape=rectangle] (y) at (1.5,-2) {output};
            \end{scope}

@@

\begin{figure}
  \centering
  \begin{subfigure}{0.5\textwidth}
    \centering
\begin{tikzpicture}
\begin{scope}[every node/.style={circle,thick,draw}]
  \node (A) at (0,0) {$x_{1}$};
    \node[dashed] (B) at (4,0) {$x_{2}$};
    \node[shape=rectangle] (U) at (2,2) {input};
    \node[shape=rectangle] (y) at (2,-2) {output};
\end{scope}

\begin{scope}[>={Stealth[black]},
              every node/.style={fill=white,circle},
              every edge/.style={draw=black, thick}]
    %\path [->] (A) edge[bend left] node {$c$} (B);
    \path [->] (B) edge node {\tiny $\gamma$} (A); 
    \path[->] (U) edge[dashed] node {\tiny $1$} (A);
    \path[->] (U) edge[dashed] node {\tiny $1$} (B);
    \path[->] (A) edge[dashed, bend right] node {\tiny $1$} (y);
\end{scope}
\begin{scope}[>={Stealth[black]},
              every edge/.style={draw=black, thick}]
    \path [->] (A) edge[loop left] node {\tiny $\lambda_{1}$} (A);
    \path [->] (B) edge[loop right] node {\tiny $\lambda_{2}$} (B);
\end{scope}

\end{tikzpicture}
    \caption{Gene Regulatory Network $A(0)$.}
  \end{subfigure}%
  \begin{subfigure}{0.5\textwidth}
    \centering
\begin{tikzpicture}
\begin{scope}[every node/.style={circle,thick,draw}]
  \node (A) at (0,0) {$x_{1}$};
    \node[dashed] (B) at (4,0) {$x_{2}$};
    \node[shape=rectangle] (U) at (2,2) {input};
    \node[shape=rectangle] (y) at (2,-2) {output};
\end{scope}

\begin{scope}[>={Stealth[black]},
              every node/.style={fill=white,circle},
              every edge/.style={draw=black, thick}]
    \path [->, sloped] (A) edge[bend left] node {\tiny $(\lambda_{2} - \lambda_{1} - \frac{\gamma}{2})$} (B);
    \path [->, sloped] (B) edge[bend left] node {\tiny $\frac{\gamma}{2}$} (A); 
    \path[->] (U) edge[dashed] node {\tiny $1$} (A);
    \path[->] (U) edge[dashed] node {\tiny $1$} (B);
    \path[->] (A) edge[dashed, bend right] node {\tiny $1$} (y);
\end{scope}
\begin{scope}[>={Stealth[black]},
              every edge/.style={draw=black, thick}]
    \path [->] (A) edge[loop left] node {\tiny $(\lambda_{1} + \frac{\gamma}{2})$} (A);
    \path [->] (B) edge[loop right] node {\tiny $(\lambda_{2}-\frac{\gamma}{2})$} (B);
\end{scope}
